---
title: "Real-Time TOC model for late-2026"
author: "Sam Struthers, B Steele & Matt Ross-ROSSyndicate"
date: "`r Sys.Date()`"
output: html_document
---

# Set Up

## Package Load

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("src/setup_libraries.R")
source("src/setup_ross_theme.R")
set.seed(123)
#set version for saving (update when changes are significant)
version = "20260715"

#### Data setup functions ####
#load grab sample data function
source("src/collation/load_grab_sample_data.R")
#load sensor data
source("src/collation/load_sensor_data.R")
#create lagged parameters
source("src/modeling/calc_lagged_parameter.R")
#create matchup dataset
source("src/modeling/create_grab_sensor_matchup.R")
# apply fdom temperature corrections
source("src/sensor_qaqc/apply_fdom_corrections.R")
##### Modeling functions #####
# function to apply sensor transformations for TOC modeling
source("src/modeling/apply_sensor_transformations_toc.R")
#function to apply training scale
source("src/modeling/apply_training_scale.R")
#function to look at xgboost feature tuning
source("src/modeling/xgboost_feature_selection.R")
# function to tune xgboost hyperparameters
source("src/modeling/xgboost_hyperparameter_tuning.R")

#### Performance and plotting functions #####

#function to plot performance of each fold
source("src/plotting/plot_tv_fold_perf.R")
#function to plot performance of each fold
source("src/plotting/plot_tvt_fold_perf.R")
#function to plot performance of each fold
source("src/plotting/plot_train_test_ensemble_perf.R")
#function to plot model timeseries
source("src/plotting/plot_model_timeseries.R")

# Load CDWR credentials to get flow data (without hitting data limits)
cdwr_creds <- read_yaml(here("creds", "CDWRCreds.yml"))$api_key

#set warnings to F
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.dim = c(12,10))
```

## Read in sensor dataset

Here, we read in the sensor data, apply min-max scaling defined in the `apply_sensor_transformations_toc()` function, and store the dataset for use later

```{r}

sensor_data <- load_sensor_data()%>% 
  filter(DT_round >= ymd("2023-08-01") & DT_round <= ymd("2026-12-31"))%>%
  mutate(DT_hourly_round = floor_date(DT_round, "hour")) %>%
  distinct(site, DT_hourly_round, parameter, .keep_all = T) %>%
  # remove parameters that are not consistently available across sites
  filter(parameter %nin% c("Chl-a ugl","TAL PC Fluorescence","ORP", "pH", "DO"))%>%
  #Remove a FC sites where there is no FDOM sensor
  filter(site %nin% c("pbr_fc", "pman_fc"))%>%
  pivot_wider(id_cols = c("site","DT_hourly_round"), names_from = "parameter", values_from = "hourly_median")%>%
  mutate(Turbidity = case_when(Turbidity < 0.1 ~ 0.1, #replace 0 turbidity with 0.1 to be able to do turbidity corrections and log transformations
                               Turbidity > 1000 ~ 1000,
                               TRUE ~ Turbidity))%>% 
  #remove rows where everything other than site and DT_round are NA
  filter(!if_all(-c(site, DT_hourly_round), is.na))%>%
  #applying FDOM temperature & turbidity correcting/standardization
  apply_fdom_corrections(fdom_col = "FDOM Fluorescence",
                         temp_col = "Temperature",
                         turb_col = "Turbidity",
                         fdom_temp_col = "FDOMc",
                         fdom_turb_col = "FDOM_turb_corr",
                         fdom_final_col = "FDOM_final_corr")

```

## Calculate additional variables on sensor data

Interaction Terms

- FDOM X Turbidity (fdom_x_turb): A proxy for turbidity interference in FDOM signal, rather than correcting for turbidity. 
- FDOM / Turbidity (fdom_turb): Another proxy measurement for FDOM and turbidity interactions
- FDOM X Specific Conductivity (fdom_x_sc): Interaction between SC and FDOM to capture changes in FDOM signal due to changes in water source (ie agricultural vs forested) that are not fully captured FDOM alone

Climate
- Seasonality (sin(doy))
- Daily flow at the watershed mouth
Adding in canyon mouth (CLAFTCCO Site) flow as a general setting for the river. Ideally we would use flow data from each site, but there are some significant spatial gaps across the watershed and we want to be sure we have a consistent flow dataset across all sites for modeling. 


```{r}

sensor_lagged <- sensor_data %>%
  mutate(year = year(DT_hourly_round)) %>%
  split(.$year) %>%
  map(., function(df){
    df %>%
      mutate(sin_doy = sin(2*pi*yday(DT_hourly_round)/365.25), 
             fdom_x_turb = FDOMc * Turbidity, 
             fdom_turb = FDOMc/Turbidity, 
             fdom_x_sc = FDOMc * `Specific Conductivity`)
  })%>%
  bind_rows()

min_date <- min(sensor_lagged$DT_hourly_round, na.rm = T)
max_date <- max(sensor_lagged$DT_hourly_round, na.rm = T)

canyon_mouth_flow_daily <- get_telemetry_ts(abbrev = "CLAFTCCO", 
                                            parameter = "DISCHRG", 
                                            start_date = as.Date(min_date)- days(1),
                                            end_date = as.Date(max_date)+ days(1),
                                            timescale = "hour",
                                            api_key = cdwr_creds)%>%
  select(abbrev, DT_hourly_round = meas_date, flow_cfs = meas_value)%>%
  mutate(DT_hourly_round = with_tz(ymd_hms(DT_hourly_round, tz = "America/Denver"), tzone = "MST"))%>%
  group_by(date = as_date(DT_hourly_round))%>%
  summarise(canyon_mouth_daily_flow_cfs = mean(flow_cfs, na.rm = T))%>%
  ungroup()


sensor_final <- sensor_lagged%>%
  mutate(date = as_date(DT_hourly_round))%>%
  left_join(canyon_mouth_flow_daily, by = "date")%>%
  select(-date, -year)%>%
  rename(DT_round = DT_hourly_round)%>%
  #not keeping the fully corrected FDOM values since this method might not be effective/correct for our sensors
  select(-FDOM_final_corr, -FDOM_turb_corr) 

```


## Read in grab sample dataset

```{r}
water_chem <- load_grab_sample_data()%>%
  #removing spring creek and box creek since they don't have good sensor data and we won't be including them in modeling
  filter(site_code %nin% c("springcreek", "boxcreek")) %>%
  #removing pbr/pman samples pre ross sensor deployment
  filter(!c(year(DT_sample) < 2025 & site_code %in% c("pbr", "pman")))%>% 
  # Ensure we have a TOC value
  filter(!is.na(TOC),
         #site_code should be in sensor dataset
         site_code %in% unique(sensor_data$site),
         # Ensure an exact DT for matchup
         !is.na(DT_sample),
         #Filter for after sensors were deployed (2023-08-01) and TOC sample collection began
         DT_sample >= ymd("2023-08-01"))%>%
  select(site = site_code, DT_sample, TOC, collector)

```
# Create Matchup

## Prep and select features 
```{r}
#We want to predict TOC
target <- "TOC"
#Sensor Data Columns we don't want to matchup
id_cols <- c("site", "DT_round", "sensor_datetime", "DT_sample", "sample_datetime","Chl-a Fluorescence", "Depth", "collector")
#Get remaining features
features <- setdiff(names(sensor_final), c(id_cols))

```

## Run matchup function 

We will use the `create_grab_sensor_matchup()` function to create the matchup dataset between grab sample chl-a data and sensor data. We will use a 3 hour acceptable time window for matching. For a "match" to be made, a row within 3 hours of the sample must have  all the parameters in `features` must have a non-NA value.

```{r}

if(file.exists(here("data", "derived", "matchup_datasets", paste0("parsed_toc_sensor_", version, ".parquet")))){
  matchup_results <- read_parquet(here("data", "derived", "matchup_datasets", paste0("parsed_toc_sensor_", version, ".parquet")))
} else {
  matchup_results <- create_grab_sensor_matchup(
    water_chem_df = water_chem,
    sensor_df = sensor_final,
    features = features,
    acceptable_time_window_hours = 3
  )
}

```

## Check Matches

```{r}

matchup_results%>%
  filter(site %in% unique(sensor_final$site))%>%
  mutate(year = year(sample_datetime))%>%
  summarise(n_missing = sum(is.na(`sensor_datetime`)),
            n_total = n(),
            perc_missing = n_missing/n_total*100, .by = year)

missing_samples <- matchup_results%>%
  mutate(year = year(sample_datetime))%>%
  group_by(year, site) %>%
  summarise(n_total = n(),
            n_missing = sum(is.na(sensor_datetime)),
            .groups = "drop")%>%
  arrange(desc(n_missing))%>%
  filter(n_missing > 0)

cat("Number of samples with no match up:\n", sum(missing_samples$n_missing), " of ", nrow(matchup_results) )
print(missing_samples, n = Inf)

#Checking specific site/years to see why there are missing matches.
matchup_results%>%
  filter(is.na(sensor_datetime) & year(sample_datetime) == 2026)%>%
  arrange(site, sample_datetime)%>%select(site, sample_datetime, collector)


### Save Matchup file

write_parquet(matchup_results,here("data", "derived", "matchup_datasets", paste0("parsed_toc_sensor_", version, ".parquet")))

```

At this point, I now have our matchup dataset and am going to see how our previous model does on the data

# Test on 2026-05-18 model 

```{r}
old_version <- "20260518"
 old_full_model <-  map(1:3, ~{
    list(
      model = xgb.load(file.path("data/derived/models/TOC_realtime/in_dev", paste( "ross_only_toc_xgboost_model_fold", .x, "_", old_version, ".ubj", sep = "")))
    )
   })

 scaled_new_data <- matchup_results %>%
  # Only keep columns that would be in the complete sensor dataset
  select(any_of(id_cols), any_of(names(sensor_final)), any_of(target))%>%
  # Apply normalization 
  apply_training_scale(
    new_data = .,
    features = features,
    scaling_params = read_parquet(file = "data/derived/model_scaling_parameters/scaling_params_toc_20260518.parquet")
  )%>%
  na.omit()%>%
     rename(id = site)
 
 test_set_2026 <- scaled_new_data %>% filter(year(sensor_datetime) == 2026)%>%
   select(id,sensor_datetime, collector,TOC,  any_of(features))
 
 train_val <- scaled_new_data%>%
   filter(year(sensor_datetime) != 2026)%>%
   filter( id %nin% c("springcreek", "boxcreek"))%>%
   filter(!(year(sensor_datetime) == 2025 & id %in% c("chd", "sfm", "pbr", "pman", "pbd", "pfal")))%>% 
  mutate(fold_id = case_when(
    id %in% c("sfm","lbea","chd" ,"riverbluffs","salyer", "elc") ~ 1,
    id %in% c( "pbd", "penn", "joei", "archery" ) ~ 2,
    id %in% c("pfal","udall","cottonwood", "cbri", "riverbend") ~ 3
  )) 
 

 ensemble <- plot_train_test_ensemble_perf(train_val_df = train_val,
                                          test_df = test_set_2026,
                                          fold_models = map(old_full_model, "model"),
                                          target_col = "TOC",
                                          units = "mg/L",
                                          subtitle_arg = "Test Set: 3/26-6/26  Model Version: 20260518") + theme(    aspect.ratio = 1 )
 
ensemble 
# ggsave(here("docs","uclp_dss", "figs","realtime_toc_modeling","performance",paste0("model_ensemble_2026_test_w_", old_version, ".png")),
#        width = 10, height = 10, dpi = 500)

plots<- map(1:3, ~{
  plot_tvt_fold_perf(train_val, test_set_2026, fold_set = .x, fold_model = old_full_model[[.x]]$model, target_col = "TOC", units = "mg/L", subtitle_arg = "Test Set: 3/26-6/26  Model Version: 20260518")
})
# Wrap plots, force them to be square, and collect the legend
combined_plot <- wrap_plots(plots, ncol = 2, guides = "collect") &
  theme(
    legend.position = "bottom",
    aspect.ratio = 1
  )
combined_plot
# ggsave(
#   here("docs", "uclp_dss", "figs", "realtime_toc_modeling", "performance", 
#        paste0("model_all_folds_TVT_2026_test_w_", old_version, ".png")),
#   plot = combined_plot,
#   width = 14, 
#   height = 14, 
#   dpi = 500
# )
```

So looks like we still continue some of the same issues (bench at ~6.5 mg/L) and generally under predicting with some wide variation between folds. We will now add in the 2025 test set into the training and test on the 2026 dataset


# Update Model

## Create Scaling Parameters & Normalize Data

Normalization is applied based on min-max scaling from the complete sensor dataset using the `apply_training_scale()` function.

```{r}
# Apply scaling to dataset
all_matches_norm <- matchup_results %>%
  # Only keep columns that would be in the complete sensor dataset
  select(any_of(id_cols), any_of(names(sensor_final)), any_of(target))%>%
  # Apply normalization 
  apply_training_scale(
    new_data = .,
    features = features, 
    full_dataset = sensor_final,
    scaling_param_out_path = paste0("data/derived/model_scaling_parameters/scaling_params_toc_", version)
  )%>%
  na.omit()

# make distribution of all matches normalized features to check for any weird outliers or issues with scaling
all_matches_norm %>%
  pivot_longer(cols = any_of(features), names_to = "feature", values_to = "value") %>%
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30, fill = "steelblue") +
  facet_wrap(~feature, scales = "free") +
  labs(title = "Distribution of Normalized Features in Matchup Dataset", x = "Normalized Value", y = "Count") +
  theme_bw()
  
```

### View correlations between features and target (TOC)
Most products with FDOM in them are going to have higher correlation with TOC, as well as seasonal (sin_doy) and flow related features. We want to be sure that we don't put too many of these in the model so we are not splitting feature importance. 

```{r}

# Combine training/validation and testing data
all_matches_norm%>%
  mutate(
    watershed = ifelse(
      site %in% c("joei", "cbri", "chd", "pfal", "pbr", "sfm", "pman", "pbd", "penn", "lbea"),
      "UpperPoudre", 
      "LowerPoudre"
    )
  )%>%
    pivot_longer(cols = any_of(features), names_to = "feature", values_to = "value")%>%
  ggplot(., aes(x = value, y = TOC, shape = collector, color = watershed)) +
    geom_point() +
    stat_cor(aes(x = value, y = TOC), inherit.aes = F, label.x.npc  = "middle")+
    labs(x = "Value", y = "TOC (mg/L)", shape = "Collector", color = "Watershed") +
    facet_wrap(~feature, scales = "free") +
    ROSS_theme

```

## Create Holdout Dataset

```{r}
# Trim out any known bad data or sites we don't want to include in modeling
all_matches_trimmed <- all_matches_norm%>%
  rename(id = site)%>%
filter( id %nin% c("springcreek", "boxcreek"))
# Create testing set from 2025 data at specific sites
testing <- all_matches_trimmed %>%
  filter((year(sensor_datetime) == 2026))
# Create training/validation set
train_val <- all_matches_trimmed %>%
  anti_join(testing)%>%
  select(id,sensor_datetime, collector,TOC,  any_of(features))

test_prop <- round((nrow(testing)/(nrow(testing) +  nrow(train_val)))*100, digits = 2)
# Calculate proportions
print(paste0("Percent of full dataset in test: ", test_prop, "%"))
print(paste0("Percent of full dataset in Training/Validation: ", 100-test_prop, "%"))

```

Currently this is a bit lower than we would like (~20%) but with the remainder of 2026 samples to be collected (bi-weekly for ~5 months at 6 sites and monthly for FC sampling at 4 sites = +70 samples)

## Visualize TOC distribution in training/validation vs testing sets

```{r}

full_data <- train_val%>%
  mutate(group = "Train/Val")%>%
  bind_rows(
    testing%>%
      mutate(group = "Testing")
  )

  full_data%>%
    ggplot(., aes(x = TOC, fill = group)) +
  geom_density(alpha = 0.5) +
  labs(
    title = paste0("Training/Validation Set: ", nrow(train_val), " samples              Testing Set: ", nrow(testing), " samples"),
    subtitle = paste0("Training TOC Range: ", round(min(train_val$TOC, na.rm = TRUE),2), " - ", round(max(train_val$TOC, na.rm = TRUE),2) , " mg/L (Median: ", round(median(train_val$TOC), 2), ")\nTesting TOC Range: ", round(min(testing$TOC, na.rm = TRUE),2), " - ", round(max(testing$TOC, na.rm = TRUE),2), " mg/L (Median: ", round(median(testing$TOC), 2),")"),
    x = "TOC (mg/L)", 
    fill = "Model Set"
  ) +
  #facet_wrap(~group, ncol = 1)+
  ROSS_theme+
  scale_fill_manual(values = c("Train/Val" = "#256BF5", "Testing" = "#E70870")) +
  theme(legend.position = "bottom", 
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 16),
        legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 16),
        plot.title = element_text(size = 18, face = "bold"),
        plot.subtitle = element_text(size = 16))

  
ggsave(here("docs", "uclp_dss", "figs", "paper_figs", paste0("toc_train_val_test_toc_distribution_", version, ".png")), width = 12, height = 8)
```

Currently a bit skewed on our test set to higher end TOC values, which makes sense since this data is March - June. Hopefully our distributions will overlap a bit more as we collect more samples

## Set up model folds
Since we have relatively little data, we are going to use three CV folds. Create 3 separate folds and 3 models, with one fold used for validation on each model. Generally we are looking for:
- Each fold to have ~30-35% of the total data in validation
- Each validation set to have a similar TOC distribution as its training set (median, range)
- Each should have a mix of Upper Mainstem (SFM, PFAL, PBD, ), Upper Trib (CBRI, CHD, LBEA, PENN) & Lower Mainstem (salyer, udall, riverbend, elc, archery, riverbluffs)

```{r}
val_set_1 <- c("sfm","elc","udall" ,"salyer")

val_set_2 <- c( "pbd", "lbea", "riverbend", "pbr" )

val_set_3 <- c("chd", "archery", "joei",  "pman")

val_set_4 <- c("pfal","penn", "cottonwood","riverbluffs", "cbri")
#Double check we covered all sites
#setdiff(unique(train_val$id), c(val_set_1, val_set_2, val_set_3, val_set_4))

# Function to create plots for a given fold
map(1:4, function(fold_id) {
  val_set_name <- paste0("val_set_", fold_id)
  val_set <- train_val %>% filter(id %in% get(val_set_name))
  train_set <- train_val %>% filter(id %nin% get(val_set_name))
  
  prop_val <- round(nrow(val_set)/nrow(train_val), 3) * 100
  train_med <- round(median(train_set$TOC, na.rm = TRUE), 1)
  val_med <- round(median(val_set$TOC, na.rm = TRUE), 1)
  
  # Custom text sizing theme layer
  text_size_theme <- theme(
    axis.title   = element_text(size = 18), 
    axis.text    = element_text(size = 18),
    legend.title = element_text(size = 18),
    legend.text  = element_text(size = 18), 
    plot.title = element_text(size = 20, face = "bold"),
    plot.subtitle = element_text(size = 18)
  )
  
  # Histogram
  hist <- ggplot(train_set, aes(x = TOC)) +
    geom_density(fill = "#256BF5", alpha = 0.5) +
    geom_density(data = val_set, aes(x = TOC), fill = "#E70870", alpha = 0.5) +
    labs(
      title = paste0("Fold Set ", fold_id, ": ", nrow(val_set), " samples (", prop_val, "%) in validation set"),
      subtitle = paste0("Validation Sites: ", paste(unique(val_set$id), collapse = ", ")),
      x = "TOC (mg/L)"
    ) +
    ROSS_theme +
    text_size_theme
  
  # Boxplot
  box <- ggplot(
    train_set %>% mutate(set = "Train") %>% 
      bind_rows(val_set %>% mutate(set = "Validation")) %>% 
      mutate(set = factor(set, levels = c("Train", "Validation"))),
    aes(x = set, y = TOC, fill = set)
  ) +
    geom_boxplot() +
    scale_fill_manual(values = c("Train" = "#256BF5", "Validation" = "#E70870")) +
    labs(
      title = paste0("Train Median TOC: ", train_med, "\nVal Median TOC: ", val_med),
      x = "",
      y = "TOC (mg/L)"
    ) +
    ROSS_theme +
    text_size_theme+
    theme(legend.position = "none")
  
  # Combine plots for this fold
  ggarrange(hist, box, nrow = 1)
}) %>%
  # Plot all folds together
  ggarrange(plotlist = ., ncol = 1)

#ggsave(here("docs", "uclp_dss", "figs", "paper_figs", paste0("toc_train_val_fold_distributions_", version, ".png")), width = 16, height = 12)

#Update train_val with a fold_id column based on val_sets
train_val <- train_val %>%
  mutate(fold_id = case_when(
    id %in% val_set_1 ~ 1,
    id %in% val_set_2 ~ 2,
    id %in% val_set_3 ~ 3, 
    id %in% val_set_4 ~ 4
  ))

```

# Model Training

For our model training we will first do some feature tuning to see which features are most useful for predicting TOC. We will then do hyperparameter tuning on the selected features to optimize model performance.

## Feature Tuning 

### Pruning Features with high correlations 

We want to avoid using too many highly correlated features in the model, as this can lead to overfitting and will split model importance. We will look at the correlation matrix and remove some of the more redundant features.

```{r}
# get a correlation matrix across all features
corr_matrix <- train_val %>% 
  select(all_of(features)) %>% 
  cor(use = "pairwise.complete.obs")

# we just want half the matrix, so recode those in the lower triangle to NA
corr_matrix[lower.tri(corr_matrix, diag = TRUE)] <- NA

# Convert to a tidy data frame and filter for those with correlation > 0.9
high_corr_pairs <- corr_matrix %>%
  as.data.frame() %>%
  rownames_to_column(var = "feature1") %>%
  pivot_longer(-feature1, names_to = "feature2", values_to = "correlation") %>%
  filter(!is.na(correlation), abs(correlation) > 0.9) %>%
  arrange(desc(abs(correlation)))

# look at the list and make some educated decisions about features
high_corr_pairs

fewer_features <- features[features %nin% c("FDOM Fluorescence")]

# do the correlation matrix again to make sure that is okay
corr_matrix <- train_val %>% 
  select(all_of(fewer_features)) %>% 
  cor(use = "pairwise.complete.obs")

corr_matrix[lower.tri(corr_matrix, diag = TRUE)] <- NA

# Convert to a tidy data frame and filter
high_corr_pairs <- corr_matrix %>%
  as.data.frame() %>%
  rownames_to_column(var = "feature1") %>%
  pivot_longer(-feature1, names_to = "feature2", values_to = "correlation") %>%
  filter(!is.na(correlation), abs(correlation) > 0.85) %>%
  arrange(desc(abs(correlation)))

high_corr_pairs

```

`high_corr_pairs` has features we should be cautious of using together. We will try to avoid using more than one of these in the same model.

### Testing Features

First, we will start with our base model (live sensor data only) and then test adding in other features to see if they improve performance. We will use the default hyperparameters below for all models to keep things consistent. Once we have tested all the features together, we will want to trim it down a bit so we are not using too many highly correlated features or overfitting the model due to lack of data

#### Run across folds testing different features to see which improve performance

Here I run the `xgboost_feature_selection()` function which trains a xgboost model with the specified features and default hyperparameters. It returns feature importance, shap values, and best iteration info for each fold. We will then look at feature importance and shap values to see which features are most useful, then trim features back to reduce our correlative and overfitting issues. 

#### Weighting functions

Because high TOC values are more important for water treatment but are relatively few in the dataset, we can use a weighting function to increase the "penalty" for getting these values wrong. Historically, we have the "TOC_high_weights" function below but I believe this might be creating part of our benching problem as it creates an antificial weight at ~4-6 mg/L. Instead I am testing out an exponential weighting function ("toc_high_weights_exp"). The weights are plotted below:

```{r}
# Weighting function for TOC
toc_high_weights <-  function(df) {
  ifelse(df$TOC <= 4,1,                  # flat weight for low TOC
         1 + 3 *log1p(df$TOC - 4))  # grows slowly, no cap needed
}

toc_high_weights_exp <- function(df) {
  ifelse(df$TOC <= 4,1,                  # flat weight for low TOC
         exp(0.5 * (df$TOC - 4)))  # exponential growth for higher TOC
}
toc_data <- tibble(TOC = seq(0, 11, by = 0.1)) %>%
  mutate(
    `Log1p Growth (Slow)` = toc_high_weights(.),
    `Exponential Growth (Fast)` = toc_high_weights_exp(.)
  ) %>%
  # Pivot longer to make plotting with ggplot easy
  tidyr::pivot_longer(
    cols = -TOC, 
    names_to = "Weight_Function", 
    values_to = "Weight"
  )

# Plot the curves
ggplot(toc_data, aes(x = TOC, y = Weight, color = Weight_Function)) +
  geom_line(linewidth = 1.2) +
  # Add a vertical reference line where the weight functions split (TOC = 4)
  geom_vline(xintercept = 4, linetype = "dashed", color = "grey50") +
  annotate(
    "text", x = 3.8, y = 15, label = "TOC Threshold (4 mg/L)", 
    angle = 90, vjust = -0.5, color = "grey40", size = 3.5
  ) +
  labs(
    title = "Comparison of TOC Weighting Functions",
    subtitle = "Comparing Log1p vs. Exponential growth models above 4 mg/L",
    x = "Measured TOC (mg/L)",
    y = "Calculated Weight Value",
    color = "Function Type"
  ) +
  scale_color_manual(values = c("Exponential Growth (Fast)" = "firebrick", "Log1p Growth (Slow)" = "steelblue")) +
  ROSS_theme +
  theme(legend.position = "bottom")
```

#### Hyperparameters 

```{r}
#raw sensor features
general_features <- c("FDOMc","Turbidity","Specific Conductivity", "Temperature")

hyper_params <- list(
  objective        = "reg:squarederror",
  eval_metric      = "rmse",
  eta              = 0.05,
  gamma            = 0.8,
  alpha            = 0,
  lambda           = 1,
  max_depth        = 3,
  subsample        = 0.5,
  colsample_bytree = 0.5,
  min_child_weight  = 2, 
  nthread = 1)

#setup function
feature_results <- map(
  #list of different features we want to test
  list(
    gen_feat = general_features, 
    fewer_features = c(fewer_features), # this is the set of features we are testing that have been pruned based on correlation and domain knowledge, but still includes flow
    # removing FDOM_turb and turb median 6 hour
    fewer_features_no_fdom_turb = c(setdiff(fewer_features, c("fdom_turb"))), 
    fewer_features_no_doy = c(setdiff(fewer_features, c("fdom_turb", "sin_doy"))),
    fewest_features = c(setdiff(fewer_features, c("fdom_turb", "sin_doy"))), 
    gen_fdom_turb_flow  = c(general_features, "fdom_x_turb","canyon_mouth_daily_flow_cfs")
  ),
  #Setting function parameters
  ~xgboost_feature_selection(
    train_val = train_val,
    features_to_test = .x,
    target = "TOC",
    default_hyper_params =  hyper_params,
    weight_fun = toc_high_weights_exp
  ))
```

#### Feature Set Results

##### Weighted RMSE

We compare across folds to see if there is big differences in folds (ie unbalanced data) and also compare each feature set against "base" model of general features (sensors). Looking for features that make meaningful improvements in RMSE across all folds. If our RMSE improves but that is related to only a single fold, we may want to reevaluate the feature set as it is likely overfitting to a given site(s). 
Keep in mind that this is weighted RMSE so it may be higher than unweighted RMSE you see in other places.

```{r}
# Highlight differences between folds in rmse
map_dfr(feature_results, ~{
  # Compute feature set description
  unique_features <- setdiff(.x$features, general_features)
  final_features <- c("general_features", unique_features)
  feature_desc <- paste(final_features, collapse = ", ")
  
  # Take best_msg tibble and rename eval_rmse by fold
  .x$best_msg %>%
    select(fold_id, eval_rmse) %>%
    mutate(fold_col = paste0("rmse_fold_", fold_id)) %>%
    select(-fold_id) %>%
    mutate(eval_rmse = round(eval_rmse, 2)) %>%
    pivot_wider(names_from = fold_col, values_from = eval_rmse) %>%
    mutate(
      features = feature_desc,
      mean_rmse = round(mean(c_across(starts_with("rmse_fold_")), na.rm = TRUE), 2)
    )%>%
    select(features, mean_rmse, contains("rmse_fold_"))
  
}, .id = "feature_set") %>%
  #check if new feature is better than baseline
  mutate(
    baseline_rmse = mean_rmse[feature_set == "gen_feat"],
    improves_rmse = if_else(mean_rmse < baseline_rmse, round(mean_rmse - baseline_rmse, 2), 0)
  ) %>%
  select(-baseline_rmse)%>%
  arrange(improves_rmse)
```
Overall, all models with sin_doy appear to do slightly better but I am a bit concerned with using this feature since we only have two years of data which had very compareable doy - TOC patterns. This pattern may not be present in future years and so we may not want to encode it in quite yet. 

It also appears that Fold 2 performs the best overall and Fold 4 performs the worst. 

##### Feature Importance

This comes directly from the xgboost package and is based on Gain, Cover, and Frequency of splits. We summarize across folds to see which features are most important across all folds. 

```{r}
#table with all importances
importance_tables <- map(feature_results, ~{
  imp <- .x$model_importance
  
  # Identify fold columns dynamically
  gain_cols <- grep("^Gain_fold_", names(imp), value = TRUE)
  cover_cols <- grep("^Cover_fold_", names(imp), value = TRUE)
  freq_cols <- grep("^Frequency_fold_", names(imp), value = TRUE)
  
  imp %>%
    mutate(
      mean_gain = rowMeans(across(all_of(gain_cols)), na.rm = TRUE),
      mean_cover = rowMeans(across(all_of(cover_cols)), na.rm = TRUE),
      mean_frequency = rowMeans(across(all_of(freq_cols)), na.rm = TRUE)
    ) %>%
    mutate(new_feat = ifelse(Feature %in% general_features, "", "*")) %>%
    select(Feature, new_feat, mean_gain, mean_cover, mean_frequency) %>%
    
    arrange(desc(mean_gain))
}, .id = "feature_set")
importance_tables

#Get features with less than 0.02 gain on average across folds in each model run to consider removing
cat("Features with mean gain < 0.02 in :\n")
iwalk(importance_tables, ~{
  low_gain_features <- .x %>%
    filter(mean_gain < 0.02) %>%
    pull(Feature)
  if(length(low_gain_features) == 0){
    #print the name of the list and say no low gain features
    cat(paste0(.y, ": No low gain features\n"))
  } else {
    cat(.y, paste(low_gain_features, collapse = ", "), "\n")
  }
})

```


##### Shapley Values

Let's look at training and validation shap values for features across all folds. This gives us a sense of which features are most important in the model predictions. We summarize across folds to see which features are most important across all folds. In general we would expect features to have similar importance in training and validation, if a feature has high importance in training but low in validation it may be overfitting to the training data and we may want to re-evaluate. 

```{r}
#table with all shap values
shap_summary <- map_dfr(feature_results, ~{
  shap <- .x$shap_values
  
  val_shap_cols <- grep("^val_mean_abs_shap_fold_", names(shap), value = TRUE)
  train_shap_cols <- grep("^train_mean_abs_shap_fold_", names(shap), value = TRUE)
  
  shap %>%
    mutate(
      mean_val_shap = round(rowMeans(across(all_of(val_shap_cols)), na.rm = TRUE), 3),
      mean_train_shap = round(rowMeans(across(all_of(train_shap_cols)), na.rm = TRUE), 3)
    ) %>%
    mutate(val_train_diff = mean_val_shap - mean_train_shap) %>%
    select(Feature, mean_val_shap, mean_train_shap, val_train_diff)
}, .id = "feature_set")

shap_long <- shap_summary %>%
  pivot_longer(
    cols = starts_with("mean_"),
    names_to = "type",
    values_to = "mean_shap"
  ) %>%
  mutate(
    type = recode(type,
                  mean_val_shap = "Validation",
                  mean_train_shap = "Training"),
    feature_set = factor(feature_set, levels = unique(feature_set))
  ) %>%
  mutate(Feature = factor(Feature, levels = c( rev(feature_results$fewer_features$features),"Sensor_Turb","Temp")))

ggplot(shap_long, aes(x = feature_set, y = Feature, fill = mean_shap)) +
  geom_tile(color = "white", linewidth = 0.4) +
  scale_fill_viridis_c(
    option = "magma",
    limits = c(0, 1),
    name = "Mean |SHAP|",
    direction = -1  # reverses so dark = low, bright = high (optional)
  )+
  facet_wrap(~type) +
  labs(
    x = "Feature Set",
    y = "Feature",
    title = "Mean SHAP Values by Feature and Feature Set"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank(),
    strip.text = element_text(face = "bold")
  )

```

### TV performance of best feature set

After trimming down a couple of features (minimal improvement/low shap values), we should take a quick look at model performance in TV space with our base model (default hyperparameters)

```{r}

gen_feat_models <- feature_results$gen_feat$model

feature_tv_plots_gen_feat <- map(1:4, ~plot_tv_fold_perf(fold_set = .x,
                                                train_val_df = train_val,
                                                fold_model = gen_feat_models[[.x]], # a list where each element is a model for that fold
                                                target_col = "TOC", units = "mg/L")
)
ft_tv_plot_gen <- wrap_plots(feature_tv_plots_gen_feat, ncol = 2, guides = "collect")&
  theme(legend.position = "bottom", aspect.ratio = 1)
ft_tv_plot_gen

best_models <- feature_results$gen_fdom_turb$model

feature_tv_plots <- map(1:4, ~plot_tv_fold_perf(fold_set = .x,
                                                train_val_df = train_val,
                                                fold_model = best_models[[.x]], # a list where each element is a model for that fold
                                                target_col = "TOC", units = "mg/L")
)
ft_tv_plot<- wrap_plots(feature_tv_plots, ncol = 2, guides = "collect")&
  theme(legend.position = "bottom", aspect.ratio = 1)

ft_tv_plot

combo <- wrap_plots(
  ft_tv_plot_gen, 
  ft_tv_plot,
  ncol = 2,
  guides = "collect") &
  theme(legend.position = "bottom")

  combo +
    plot_annotation(
    title = "Model Performance Comparison: Default Sensor (left) vs gen_fdom_turb Model (right)"
  )
```


## Hyperparameter Tuning
Now that we have a feature set, we can do some hyperparameter tuning to see if we can improve model performance. This will take a while to run, so we will save the results and load them in later.

### Setup

```{r}

features <- xgb.importance(model = best_models[[1]])$Feature # get features from model to use in tuning)

# Reduce tune grid for reruns (ie. General best params)
tune_grid = expand.grid(
  nrounds = 1000,
  max_depth = c(2, 3, 4),
  eta = c( 0.01,0.05, 0.1),
  gamma = c(0.8),
  lambda = c(1,10),
  alpha = c(0, 1),
  colsample_bytree = c(0.5,0.8),
  min_child_weight = c(2, 4),
  subsample = c(0.5,0.8), 
  nthread = 1
)

```

### Run hyperparameter tuning

```{r}
set.seed(123)
hyperp_results <- xgboost_hyperparameter_tuning(
  train_val_df = train_val %>% select(TOC,fold_id, any_of(features)), 
  feature_cols = features,
  weights = toc_high_weights_exp,
  tune_grid = tune_grid,
  target_col = "TOC",
  units = "mg/L")

```

### Choosing best grids

### Fold 1

```{r}
hyperp_results[[1]]$perf_dist
hyperp_results[[1]]$eval_plot
hyperp_results[[1]]$tv_plot

 best_fold_1_grid <- 29
```
Chose the grid `best_fold_1_grid` for fold 1 because it has one of the lowest validation RMSEs and it's RMSE is most closely related to it's val RMSE in the eval plot. Overall there was minimal difference between grids. Some minor benching at the lower and upper end. 

### Fold 2

```{r}
hyperp_results[[2]]$perf_dist
hyperp_results[[2]]$eval_plot
hyperp_results[[2]]$tv_plot

best_fold_2_grid <- 220

```
Chose the grid `best_fold_2_grid` for fold 2 because the validation RMSEs are pretty similar and eval plots look similar across folds. 

### Fold 3

```{r}
hyperp_results[[3]]$perf_dist
hyperp_results[[3]]$eval_plot
hyperp_results[[3]]$tv_plot

best_fold_3_grid <- 15

```

Chose the grid `best_fold_3_grid` for fold 3 because the validation RMSEs are pretty similar. We do have lower TOC values in our validation set, particularly driven by JOEI's samples. 

### Fold 4

```{r}
hyperp_results[[4]]$perf_dist
hyperp_results[[4]]$eval_plot
hyperp_results[[4]]$tv_plot

best_fold_4_grid <- 34

```

Chose the grid `best_fold_4_grid` for fold 4 because RMSEs are consistent, and the validation points hug the 1:1 line across TOC levels whereas other models appear to deviated to underpredict. The eval plots appear to show that we are finding local minimums in the validation data. 

```{r}

full_model <- map(1:4, function(i) {
  
  best_grid <- paste0("grid_", get(glue("best_fold_{i}_grid")))
  model_info <- hyperp_results[[i]]$grid_results[[best_grid]]
  list(
    model = model_info$model,
    perf = model_info$perf,
    params = model_info$params,
    training_data = train_val %>% filter(fold_id != i),
    validation_data = train_val %>% filter(fold_id == i)
  )
})

```

### Checking Results


#### Feature Importance

```{r}
importance_list <- lapply(seq_along(full_model), function(i) {
  fold_model <- full_model[[i]]$model
  features <- xgb.importance(model = fold_model)$Feature
  importance_matrix <- xgb.importance(feature_names = features, model = fold_model)
  importance_df <- as.data.frame(importance_matrix)
  importance_df$Fold <- paste("Fold", i)
  return(importance_df)
})

print(importance_list)
```

Very clear that FDOMc and fdom_x_turb are consistently the most important features and then there is a mix of importances between other features

#### Shapley Importance

```{r}
# Function to plot SHAP values for a given fold
plot_shap <- function(fold_set, training_data, fold_model) {
  #get features from model
  features <- xgb.importance(model = fold_model)$Feature
  # Predictions on validation/training set
  dtrain <- as.matrix(training_data[, features])
  
  shap_values <- shap.values(xgb_model = fold_model, X_train = dtrain)
  
  # Long-format SHAP values
  shap_long <- shap.prep(
    shap_contrib = shap_values$shap_score,
    X_train = dtrain
  )
  #return shapley plot
  shap.plot.summary(shap_long)
}

map(1:4, ~ plot_shap(
  fold_set = .x,
  fold_model = full_model[[.x]]$model,
  training_data = full_model[[.x]]$training_data
))%>%
  wrap_plots(., ncol = 1, axes = "collect", guides = "collect") &
  theme(legend.position = "bottom")

```
All the shap values appear to agree with our expectation that higher FDOM == higher TOC. I am a bit surprised to see that higher canyon mouth Q in some folds results in lower SHAP values but has follows that higher Q == higher TOC in other folds. 

### Plots of TV Performance
Now let's look at our performance in TV space.

```{r}

hyperparam_tv_plots <- map(1:4, ~plot_tv_fold_perf(fold_set = .x,
                                                   train_val_df = train_val,
                                                   fold_model = full_model[[.x]]$model, # a list where each element is a model for that fold
                                                   target_col = "TOC", units = "mg/L")
)

hp_tv_plot<- wrap_plots(hyperparam_tv_plots, ncol = 2, guides = "collect")&
  theme(legend.position = "bottom", aspect.ratio = 1)
hp_tv_plot
ggsave(here(paste0("docs/uclp_dss/figs/realtime_toc_modeling/performance/toc_model_fold_validation_performance_", version, ".png")), width = 14, height = 10)

```

#### Against base hyperparameters

We should hope that all our tuning efforts have improved the model!! 

```{r}
# Compare best hyperparameter model to default hyperparameter model
final_plot <- wrap_plots(
  ft_tv_plot,
  hp_tv_plot,
  ncol = 2,
  guides = "collect"
) &
  theme(legend.position = "bottom")
final_plot <- final_plot + 
  plot_annotation(
    title = "Model Performance Comparison: Default (left) vs Tuned Hyperparameters (right)"
  )

final_plot

```

Looking at the plots, we can see that the tuned hyperparameter model has improved our results. It appears that fold 2 and 3 has the worst performance which is likely due to the distribution of the validation set being the widest and contains the sites (JOEI: lowest TOC). We will save this model and we generally see better performance and similar performance across all folds. 

## Save as "in_dev" and remove unneeded objects

If we want to test additional ideas for bias correction or look at model performance on timeseries data, we can just run this script until `model_training` and then load in the "in_dev" model file to avoid having to rerun all the tuning again.

```{r}
#commented out so we don't overwrite existing model files

path <- here("data", "derived", "models","TOC_realtime", "in_dev")

xgb.save(full_model[[1]]$model, fname = file.path(path, paste0("ross_only_toc_xgboost_model_fold1_", version, ".ubj")))
xgb.save(full_model[[2]]$model, fname = file.path(path, paste0("ross_only_toc_xgboost_model_fold2_", version, ".ubj")))
xgb.save(full_model[[3]]$model, fname = file.path(path, paste0("ross_only_toc_xgboost_model_fold3_", version, ".ubj")))
xgb.save(full_model[[4]]$model, fname = file.path(path, paste0("ross_only_toc_xgboost_model_fold4_", version, ".ubj")))

#if we want to reload model into full model in the $model slot for each fold index
full_model <- 1:4 %>%
  map(~{
    list(
      model = xgb.load(file.path(path, paste( "ross_only_toc_xgboost_model_fold", .x, "_", version, ".ubj", sep = ""))),
      training_data = train_val %>% filter(fold_id != .x),
      validation_data = train_val %>% filter(fold_id == .x)
    )
  })

rm(list = setdiff(ls(), c("train_val", "testing", "full_model","version","id_cols",
                          "sensor_final",
                          #keeping functions around
                          "apply_training_scale","plot_tvt_fold_perf","plot_train_test_ensemble_perf",
                          "load_grab_sample_data", "plot_model_timeseries" )) )
```

# Evaluate on Test Set

We should only do this once we feel good about our model and are ready to see how it performs on the unseen test set. This should only be run after big changes to the model or if we get new data (ie The model should be in a "final/working" version)

## Visualize Train Val Test

### Individual folds

```{r}
plots <- map(1:4, ~{
  plot_tvt_fold_perf(
    train_val, 
    testing, 
    fold_set = .x, 
    fold_model = full_model[[.x]]$model, 
    target_col = "TOC", 
    units = "mg/L", 
    subtitle_arg = "Model Version: 20260715"
  )
})

plots

walk(1:4, ~{
  ggsave(
    filename = here("docs", "uclp_dss", "figs", "realtime_toc_modeling", "performance", paste0("ross_only_model_fold_", .x, ".png")),
    plot = plots[[.x]], 
    width = 10, 
    height = 6, 
    dpi = 500
  )
})

```

### All folds together

```{r}
# Single plot with all folds
wrap_plots(plots, ncol = 2, guides = "collect") &
  theme(legend.position = "bottom", aspect.ratio = 1)
ggsave(here("docs","uclp_dss", "figs","realtime_toc_modeling","performance",paste0("ross_only_model_all_folds_TVT", version, ".png")),
       width = 14, height = 16, dpi = 500)
```

In general, our RMSE is a bit higher than in previous versions. Interestingly, our Folds 3 & 4 have the worst performance (which was not the case in the validation set) so we may be slightly overfit to historical conditions. Additionally, our inclusion of lower CLP sites is likely not improving our performance here since the water sources are very different. We do see pretty consistent shape of predictions between the validation and test set which is good though. Since our test set is skewed a bit higher, it does make sense that our RMSE is a bit higher as we generally perform better at lower TOC. 

### Ensemble predictions

```{r}
# Plot ensemble
ensemble <- plot_train_test_ensemble_perf(train_val_df = train_val,
                                          test_df = testing,
                                          fold_models = map(full_model, "model"),
                                          target_col = "TOC",
                                          units = "mg/L",
                                          subtitle_arg = "Trained on: 2023-2025, Test: Sp2026, Model Version: 20260715")+theme(aspect.ratio = 1)
plot(ensemble)
ggsave(here("docs","uclp_dss", "figs", "realtime_toc_modeling","performance",paste0("ross_only_model_ensemble", version, ".png")), plot = ensemble, width = 12, height = 12, dpi = 500)
```

In ensemble space, our RMSE improves slightly and in general our predictions are closer to the 1:1 line than our previous model version. We also have generally smaller prediction ranges than in previous versions. 


# Timeseries Space Performance

## Prep sensor timeseries & water chem

```{r}
#loading in all sensor data and applying training scaling
feats <- xgb.importance(full_model[[1]]$model)$Feature

sensor_ts <- apply_training_scale(new_data = sensor_final, 
                                  full_dataset = sensor_final, 
                                  features = feats )%>%
  select(site, DT_round, any_of(feats))

water_chem <- load_grab_sample_data()

time_series_dir <- here("docs", "uclp_dss", "figs", "realtime_toc_modeling", "timeseries")
```

## Example 1: CHD 2026

```{r}

chd_plot <- plot_model_timeseries(model_input_df = sensor_ts, 
                                  water_chem_df = water_chem, 
                                  summary_interval = "6 hour",
                                  site_sel = "chd", 
                                  site_title = "Chambers Lake outflow",
                                  model = map(full_model, "model"),
                                  method = "Ensemble",
                                  target_col = "TOC",
                                  units = "mg/L",
                                  start_DT = "2026-05-01 12:00",
                                  end_DT = "2026-07-05 00:00",
                                  subtitle_arg = "v20260715", 
                                  plot_model_input = T)


ggarrange(chd_plot$model_plot, chd_plot$model_input_plot,ncol = 2)

ggsave(here(time_series_dir, paste0("chambers_outflow_20260501_20260705_toc", version,".png")),
       plot = chd_plot$model_plot,
       width = 12,
       height = 8,
       dpi = 300)

ggsave(here(time_series_dir, paste0("chambers_outflow_20260501_20260705_sensor.png")),
       plot = chd_plot$model_input_plot,
      width = 10,
      height = 10,
       dpi = 300)

```


## Example 2: PFAL 2026

```{r}

pfal_plot <- plot_model_timeseries(model_input_df = sensor_ts, 
                                   water_chem_df = water_chem%>%filter(collector == "ROSS"), 
                                   summary_interval = "6 hour",
                                   site_sel = "pfal", 
                                   site_title = "Poudre @ Poudre Falls",
                                   model = map(full_model, "model"),
                                   method = "Ensemble",
                                   target_col = "TOC",
                                   units = "mg/L",
                                   start_DT = "2026-03-10 00:00",
                                   end_DT = "2026-07-10 00:00",
                                   subtitle_arg = "CLP Samples Only", 
                                   plot_model_input = T)

ggarrange(pfal_plot$model_plot, pfal_plot$model_input_plot,ncol = 2)


 ggsave(here(time_series_dir, paste0("pfal_20260310_20260710_toc_", version, ".png")),
        plot = pfal_plot$model_plot,
        width = 12,
        height = 8,
        dpi = 300)


 ggsave(here(time_series_dir, paste0("pfal_20260310_20260710_sensor.png")),
        plot = pfal_plot$model_input_plot,
        width = 10,
        height = 10,
       dpi = 300)

```

## Example 3: PBR

```{r}


pbr_plot <- plot_model_timeseries(model_input_df = sensor_ts, 
                                   water_chem_df = water_chem, 
                                  summary_interval = "4 hour",
                                  site_sel = "pbr", 
                                  site_title = "Poudre @ Indian Meadows",
                                  model = map(full_model, "model"),
                                  method = "Ensemble",
                                  target_col = "TOC",
                                  units = "mg/L",
                                  start_DT = "2026-03-01 00:00",
                                  end_DT = "2026-07-08 00:00",
                                  subtitle_arg = "CLP Samples Only", 
                                  plot_model_input = T)

 ggarrange(pbr_plot$model_plot, pbr_plot$model_input_plot,ncol = 2)


 ggsave(here(time_series_dir, paste0("pbr_20260301_20260708_toc_", version,".png")),
       plot = pbr_plot$model_plot,
       width = 12,
       height = 8,
       dpi = 300)

 ggsave(here(time_series_dir, paste0("pbr_20260301_20260708_sensor.png")),
        plot = pbr_plot$model_input_plot,
        width = 10,
        height = 10,
        dpi = 300)
 
```

## Example 4: SFM 2026

```{r}

sfm_plot <- plot_model_timeseries(model_input_df = sensor_ts, 
                                  water_chem_df = water_chem, 
                                  summary_interval = "4 hour",
                                  site_sel = "sfm", 
                                  site_title = "South Fork of the Poudre",
                                  model = map(full_model, "model"),
                                  method = "Ensemble",
                                  target_col = "TOC",
                                  units = "mg/L",
                                  start_DT = "2026-03-01 00:00",
                                  end_DT = "2026-07-01 00:00",
                                  subtitle_arg = "CLP Samples Only", 
                                  plot_model_input = T)

ggarrange(sfm_plot$model_plot, sfm_plot$model_input_plot,ncol = 2)

ggsave(here(time_series_dir, paste0("sfm_20260301_20260701_toc_",version, ".png")),
       plot = sfm_plot$model_plot,
       width = 12,
       height = 8,
       dpi = 300)

ggsave(here(time_series_dir, paste0("sfm_20260301_20260701_sensor.png")),
       plot = sfm_plot$model_input_plot,
       width = 10,
       height = 10,
       dpi = 300)

```



## Example 4: PMAN 2026

```{r}

pman_plot <- plot_model_timeseries(model_input_df = sensor_ts, 
                                   water_chem_df = water_chem%>%filter(collector == "ROSS"), 
                                   summary_interval = "4 hour",
                                   site_sel = "pman", 
                                   site_title = "Poudre at Manners",
                                   model = map(full_model, "model"),
                                   method = "Ensemble",
                                   target_col = "TOC",
                                   units = "mg/L",
                                   start_DT = "2026-03-10 00:00",
                                   end_DT = "2026-07-01 00:00",
                                   subtitle_arg = "CLP Samples Only", 
                                   plot_model_input = T)

ggarrange(pman_plot$model_plot, pman_plot$model_input_plot,ncol = 2)



ggsave(here(time_series_dir, paste0("pman_20260301_20260701_toc_",version,".png")),
       plot = pman_plot$model_plot,
       width = 12,
       height = 8,
       dpi = 300)

ggsave(here(time_series_dir, paste0("pman_20260301_20260701_sensor.png")),
       plot = pman_plot$model_input_plot,
       width = 10,
       height = 10,
       dpi = 300)

```

## Example 5: PBD 2026

```{r}
pbd_plot <- plot_model_timeseries(model_input_df = sensor_ts, 
                                  water_chem_df = water_chem, 
                                  summary_interval = "4 hour",
                                  site_sel = "pbd", 
                                  site_title = "Poudre at Canyon Mouth",
                                  model = map(full_model, "model"),
                                  method = "Ensemble",
                                  target_col = "TOC",
                                  units = "mg/L",
                                  start_DT = "2026-03-10 00:00",
                                  end_DT = "2026-07-01 00:00",
                                  subtitle_arg = "CLP Samples Only", 
                                  plot_model_input = T)

ggarrange(pbd_plot$model_plot, pbd_plot$model_input_plot,ncol = 2)

ggsave(here(time_series_dir, paste0("pbd_20260301_20260701_toc_",version, ".png")),
       plot = pbd_plot$model_plot,
       width = 12,
       height = 8,
       dpi = 300)

ggsave(here(time_series_dir, paste0("pbd_20260301_20260701_sensor.png")),
       plot = pbd_plot$model_input_plot,
       width = 10,
       height = 10,
       dpi = 300)
```

### PBD 2024

Mostly curious how we do in the previous year (this data is in the Training set)

```{r}
pbd_plot <- plot_model_timeseries(model_input_df = sensor_ts, 
                                  water_chem_df = water_chem%>%filter(collector == "ROSS"), 
                                  summary_interval = "6 hour",
                                  site_sel = "pbd", 
                                  site_title = "Poudre at Canyon Mouth",
                                  model = map(full_model, "model"),
                                  method = "Ensemble",
                                  target_col = "TOC",
                                  units = "mg/L",
                                  start_DT = "2025-04-05 00:00",
                                  end_DT = "2025-11-05 00:00",
                                  subtitle_arg = "CLP Samples Only", 
                                  plot_model_input = T)

ggarrange(pbd_plot$model_plot, pbd_plot$model_input_plot,ncol = 2)

ggsave(here(time_series_dir, paste0("pbd_20250405_20251105_toc_",version, ".png")),
       plot = pbd_plot$model_plot,
       width = 12,
       height = 8,
       dpi = 300)
```


## Example 6: Udall (Lincoln Ave) 2025

Mostly curious how we do in the lower Poudre (this site is in the Training set)

```{r}


udall_plot <- plot_model_timeseries(model_input_df = sensor_ts, 
                                    water_chem_df = water_chem%>%filter(collector == "ROSS"), 
                                    summary_interval = "6 hour",
                                    site_sel = "elc", 
                                    site_title = "Poudre at Lincoln Ave",
                                    model = map(full_model, "model"),
                                    method = "Ensemble",
                                    target_col = "TOC",
                                    units = "mg/L",
                                    start_DT = "2025-04-15 00:00",
                                    end_DT = "2025-07-05 00:00",
                                    subtitle_arg = "CLP Samples Only", 
                                    plot_model_input = T)

ggarrange(udall_plot$model_plot, udall_plot$model_input_plot,ncol = 2)

 ggsave(here(time_series_dir, paste0("udall_20250415_20250705_toc_",version,".png")),
        plot = udall_plot$model_plot,
        width = 12,
        height = 8,
        dpi = 300)

 ggsave(here(time_series_dir, paste0("udall_20250520_20250705_sensor.png")),
        plot = udall_plot$model_input_plot,
        width = 10,
        height = 10,
        dpi = 300)
```



# Save final models

We will extract just the model objects from the list to save space.

```{r}

small_models <- map(full_model, "model")

#Save xgb versions for safety
xgb.save(small_models[[1]], fname = here("data", "derived", "models","TOC_realtime", "in_dev", paste0("ross_only_toc_xgboost_model_fold1_", version, ".ubj")))

xgb.save(small_models[[2]], fname = here("data", "derived", "models","TOC_realtime", "in_dev", paste0("ross_only_toc_xgboost_model_fold2_", version, ".ubj")))

xgb.save(small_models[[3]], fname = here("data", "derived", "models","TOC_realtime", "in_dev", paste0("ross_only_toc_xgboost_model_fold3_", version, ".ubj")))

xgb.save(small_models[[4]], fname = here("data", "derived", "models","TOC_realtime", "in_dev", paste0("ross_only_toc_xgboost_model_fold4_", version, ".ubj")))

```







